Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-09-12 LaTex/paper.tex
review changes, bib cosmetics

The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual item, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate. From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of the industrial practice of a vendor of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand. This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to  fully controlled for.

@@

In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice. Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\mu = \sigma^2$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

For a first qualitative assessment, we make use of the probability integral transform, see e.g. \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed. We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution. Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

@@

In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec. \ref{sec:cdf_histo}), inverse quantile profile plots do not work on individual PDF predictions but require a statistical population. However, calculating the fractions of the inverse quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts. Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the inverse quantile profile plot can be any variable of the data set at hand.

@@

The methods discussed so far allow a detailed qualitative evaluation of PDF predictions. However, in order to also quantify the quality of the PDF predictions, a metric is required that assesses the difference between the distributions. Therefore, we want to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly. Several approaches that measure the difference between two probability distributions are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, or the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius.

@@

Since $0.5$ represents the maximum value of the first Wasserstein distance (its minimum being zero), we define an accuracy measure for our PDF predictions in the range $[0, 1]$ by:
\begin{equation}
\text{accuracy} = 1 - 2 \cdot \text{EMD}
\end{equation}

@@

\noindent
While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the last bins of CDF values close to $1$ as well as in the first bins of CDF values close to $0$. These cases are mainly due to slow moving articles with less than one item sold per store per day on average. In practice, predicting the demand for these kind of items is very challenging as they are prone to over-dispersion and large fluctuations due to the discrete nature of the sales event. In the lower part of \fig \ref{fig:cdf_demand}, the effects of these slow-sellers are reduced by excluding all samples with mean predictions lower than $1.0$. The resulting distribution for the negative binomial model indicates a slight tendency towards too broad PDF predictions, at least for some observations.

>>>>

The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive,oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual item, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate. From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of the industrial practice of a vendor of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand. This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to fully control for.

@@

In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice. Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

For a first qualitative assessment, we make use of the probability integral transform, see e.g. \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1. We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution. Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

@@

In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec. \ref{sec:cdf_histo}), inverse quantile profile plots do not work on individual PDF predictions but require a statistical population. However, calculating the fractions of the inverse quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts. Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the inverse quantile profile plot can be any variable of the data set at hand known at prediction time.

@@

The methods discussed so far allow a detailed qualitative evaluation of PDF predictions. In order to also quantify the quality of the PDF predictions, we use a metric assessing the difference between two probability distributions to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution, and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly. Several approaches that measure the difference between two probability distributions are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, or the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius.

@@

Since $0.5$ represents the maximum value of the first Wasserstein distance when comparing any distribution in the support $[0, 1]$ to a flat distribution in the same interval (its minimum being zero), we define an accuracy measure for our PDF predictions in the range $[0, 1]$ by:
\begin{equation}
\text{accuracy} = 1 - 2 \cdot \text{EMD}
\end{equation}

@@

\noindent
While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the first bins of CDF values close to $0$ (overprediction) as well as in the last bins of CDF values close to $1$ (underprediction). The first case is mainly due to a slight zero inflation of actual sales. The latter case primarily stems from slow moving articles with less than one item sold per store per day on average, what can be seen in plots c and d of \fig \ref{fig:cdf_demand}, where the spike at CDF values close to $1$ is reduced by excluding all samples with mean predictions lower than $1.0$. The resulting distribution for the negative binomial model indicates a slight tendency towards too broad PDF predictions, at least for some observations.
